Draft -- Discriminating Seagrasses From Green Macroalgae in European Intertidal areas using high resolution multispectral drone imagery -- Draft

### 2024-09-19 index.qmd
resampling saved
updates

# res_to_resample <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,95,100,200,300,400,500,600,700,800,900,1000,1500,2000,2500,3000)
res_to_resample <- c(2000)

@@

```{r gif_making}

list_pred <- list.files("Data/Predictions/Resampled",full.names = T) %>% 
  as_tibble() %>% 
  rename(path = "value") %>% 
  mutate(name = gsub(".*/","",path),
         type = case_when(str_detect(name, "_prob.tif") ~ "Prob",
                          T ~ "Pred")) %>% 
  dplyr::filter(type == "Pred") %>% 
  mutate(res =sub(".*_(\\d+)cm_.*", "\\1", name) %>% as.numeric(),
         ortho_name  = sub("_(\\d+)cm_.*", "", name)) %>% 
  arrange(ortho_name,res) %>% 
  mutate(site = case_when(str_detect(ortho_name, "Aveiro_1506_Highest") ~ "Aveiro",
                          str_detect(ortho_name, "Aveiro_boat_1706_flight_1") ~ "Boat1",
                          str_detect(ortho_name, "Aveiro_boat_1706_flight_2") ~ "Boat2",
                          str_detect(ortho_name, "NoirmoutierNorthDikeSite") ~ "NorthDike",
                          str_detect(ortho_name, "Duer_15072022_120m") ~ "Duer",
                          str_detect(ortho_name, "handfisher_Ortho") ~ "handfisher_Ortho",
                          str_detect(ortho_name, "Horse_rooster") ~ "Horse_rooster",
                          str_detect(ortho_name, "fl1_120m_handfisher_old") ~ "fl1_120m_handfisher_old",
                          str_detect(ortho_name, "fl2_120m_handfisher_old") ~ "fl2_120m_handfisher_old",
                          T ~ NA)) %>% 
  dplyr::filter(!site == "handfisher_Ortho",
                !site == "Duer")

@@

```{r resampling validation data}

>>>>

res_to_resample <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,95,100,200,300,400,500,600,700,800,900,1000,1500,2000,2500,3000)
res_to_resample <- c(2000)

@@

```{r gif_making}
#| cache: false
#| echo: false
#| warning: false
#| eval: false
list_pred <- list.files("Data/Predictions/Resampled",full.names = T) %>% 
  as_tibble() %>% 
  rename(path = "value") %>% 
  mutate(name = gsub(".*/","",path),
         type = case_when(str_detect(name, "_prob.tif") ~ "Prob",
                          T ~ "Pred")) %>% 
  dplyr::filter(type == "Pred") %>% 
  mutate(res =sub(".*_(\\d+)cm_.*", "\\1", name) %>% as.numeric(),
         ortho_name  = sub("_(\\d+)cm_.*", "", name)) %>% 
  arrange(ortho_name,res) %>% 
  mutate(site = case_when(str_detect(ortho_name, "Aveiro_1506_Highest") ~ "Aveiro",
                          str_detect(ortho_name, "Aveiro_boat_1706_flight_1") ~ "Boat1",
                          str_detect(ortho_name, "Aveiro_boat_1706_flight_2") ~ "Boat2",
                          str_detect(ortho_name, "NoirmoutierNorthDikeSite") ~ "NorthDike",
                          str_detect(ortho_name, "Duer_15072022_120m") ~ "Duer",
                          str_detect(ortho_name, "handfisher_Ortho") ~ "handfisher_Ortho",
                          str_detect(ortho_name, "Horse_rooster") ~ "Horse_rooster",
                          str_detect(ortho_name, "fl1_120m_handfisher_old") ~ "fl1_120m_handfisher_old",
                          str_detect(ortho_name, "fl2_120m_handfisher_old") ~ "fl2_120m_handfisher_old",
                          T ~ NA)) %>% 
  dplyr::filter(!site == "handfisher_Ortho",
                !site == "Duer")

@@

```{r resampling validation data}
#| cache: false
#| echo: false
#| warning: false
#| eval: false

library(MapRs)

img_list <- list.files("Data/Predictions/RAW/", pattern = ".tif", full.names = T) %>% 
  as_tibble() %>% 
  rename(path = "value") %>% 
  mutate(name = gsub(".*/","",path),
         type = case_when(str_detect(name,"Probs") ~ "Probs",
                          TRUE ~ "Pred"),
         site = case_when(str_detect(name,"Boat1") ~ "Boat1",
                          str_detect(name,"Boat2") ~ "Boat2",
                          str_detect(name,"Gafanha_high") ~ "AveiroHigh",
                          str_detect(name,"Duer") ~ "Duer",
                          str_detect(name,"Gafanha_Low") ~ "AveiroLow",
                          str_detect(name,"NorthDike") ~ "NorthDike"),
         filtered = case_when(str_detect(name,"_7") ~ T,
                              T ~ F),
         UTM_zone = case_when(str_detect(path, "NorthDike") ~ 32630,
                              str_detect(path, "Duer") ~ 32630,
                              TRUE ~ 32629)) %>% 
  dplyr::filter(site != "AveiroLow",
                type == "Pred",
                filtered == F,
                site == "NorthDike")

# res_to_resample <- c(10,15,20,25,30,35,40,45,50,55,60,65,70,75,80,85,90,95,100,200,300,400,500,600,700,800,900,1000,1500,2000,2500,3000)
res_to_resample <- c(3000)

for(i in 1:nrow(img_list)){

  img <- img_list %>% 
    dplyr::filter(path == img_list$path[i]) %>% 
    pull(path) %>% 
    rast()%>% 
    terra::project(paste0("EPSG:", img_list$UTM_zone[i]), thread = T, method = "near")

    for(ii in 1:length(res_to_resample)){

      new_res <- res_to_resample[ii]/100

    r_template <- rast(ext(img), resolution = new_res, crs = crs(img))
    r_resampled <- terra::resample(img, r_template, method="mode", threads = T)%>% 
      terra::project("EPSG:4326", thread = T, method = "near")

    if(i == 1){
      if(!dir.exists(paste0("Data/Validation/Prediction_resampling/",res_to_resample[ii],"cm"))){
       dir.create(paste0("Data/Validation/Prediction_resampling/",res_to_resample[ii],"cm"), recursive = T) 
      }
    }
    writeRaster(r_resampled, paste0("Data/Validation/Prediction_resampling/",res_to_resample[ii],"cm/",gsub(".tif",paste0("_",res_to_resample[ii],"cm.tif"),img_list$name[i])),overwrite = T)
    }
  }
```

```{r Comparison resampled data}
#| cache: false
#| echo: false
#| warning: false
#| eval: false

df_valid <- list.files("Data/Validation/Prediction_resampling",pattern = ".tif",recursive = T, full.names = T) %>% 
  as_tibble() %>% 
  rename(path = "value") %>% 
  mutate(name = gsub(".*/","",path)) %>% 
  # dplyr::filter(type == "Pred") %>% 
  mutate(res =sub(".*_(\\d+)cm.*", "\\1", name) %>% as.numeric()) %>% 
  arrange(name,res) %>% 
  mutate(site = case_when(str_detect(name, "Gafanha_high") ~ "Aveiro",
                          str_detect(name, "Boat1") ~ "Boat1",
                          str_detect(name, "Boat2") ~ "Boat2",
                          str_detect(name, "NorthDike") ~ "NorthDike",
                          str_detect(name, "Duer_high") ~ "Duer",
                          str_detect(name, "handfisher_Ortho") ~ "handfisher_Ortho",
                          T ~ NA),
         type = "validation")

df_resampled <- list.files("Data/Predictions/Resampled",full.names = T) %>% 
  as_tibble() %>% 
  rename(path = "value") %>% 
  mutate(name = gsub(".*/","",path),
         type = case_when(str_detect(name, "_prob.tif") ~ "Prob",
                          T ~ "Pred")) %>% 
  dplyr::filter(type == "Pred") %>% 
  mutate(res =sub(".*_(\\d+)cm_.*", "\\1", name) %>% as.numeric(),
         ortho_name  = sub("_(\\d+)cm_.*", "", name)) %>% 
  arrange(ortho_name,res) %>% 
  mutate(site = case_when(str_detect(ortho_name, "Aveiro_1506_Highest") ~ "Aveiro",
                          str_detect(ortho_name, "Aveiro_boat_1706_flight_1") ~ "Boat1",
                          str_detect(ortho_name, "Aveiro_boat_1706_flight_2") ~ "Boat2",
                          str_detect(ortho_name, "NoirmoutierNorthDikeSite") ~ "NorthDike",
                          str_detect(ortho_name, "Duer_15072022_120m") ~ "Duer",
                          str_detect(ortho_name, "handfisher_Ortho") ~ "handfisher_Ortho",
                          str_detect(ortho_name, "Horse_rooster") ~ "Horse_rooster",
                          str_detect(ortho_name, "fl1_120m_handfisher_old") ~ "fl1_120m_handfisher_old",
                          str_detect(ortho_name, "fl2_120m_handfisher_old") ~ "fl2_120m_handfisher_old",
                          T ~ NA)) %>% 
  dplyr::filter(!site == "handfisher_Ortho",
                !site == "Duer") %>% 
  dplyr::select(-ortho_name)

df <- df_valid %>% 
  bind_rows(df_resampled) %>% 
  dplyr::filter(site %in% c("Aveiro","Boat1","Boat2","NorthDike")) %>% 
  # dplyr::filter(site == "Aveiro") %>% 
  dplyr::select(-name) %>% 
  pivot_wider(names_from = type, values_from = path)

for(i in 1:nrow(df)){

  print(i)

  img <- rast(c(df$validation[i],df$Pred[i]))

  names(img) <- c("true","pred")

  df_img <- img %>%  
    as.data.frame() %>% 
    dplyr::filter(!true == 0,
                  !pred == 0) %>% 
    dplyr::mutate(res = df$res[i],
                  site = df$site[i])

  if(i == 1){
    df_compare <- df_img
  }else{
    df_compare <- rbind(df_compare,df_img)
  }
}

validation = df_compare %>% 
  mutate(same = case_when(pred == true ~ T,
                          T ~ F))

total <- validation %>% 
  lazy_dt() %>% 
  group_by(res,true) %>% 
  dplyr::summarise(agreed = sum(same, na.rm = T),
          n_obs = n()) %>% 
  # ungroup() %>% 
  mutate(prop = agreed/n_obs) %>% 
  as_tibble() %>% 
  dplyr::filter(true != 1) %>% 
  mutate(true_fct = as.factor(true),
         area = res*res)

total %>% 
  ggplot(aes(x = res, y = prop, group = true, color = as.factor(true)))+
  geom_line()
  # scale_x_log10()

glm_res_true <- glm(prop ~ area * true_fct , data = total) 
glm_res <- glm(prop ~ area, data = total, family = "")

summary(glm_res)

Utilities.Package::Model_Check(glm_res)

NewData<-expand_grid(area=unique(total$area),
                    true_fct=unique(total$true_fct))

pred <- predict(glm_res_true, NewData, se.fit = T)

NewData <- NewData %>% 
  mutate(fit = pred$fit*100,
         se = pred$se.fit*100, 
         res = sqrt(area)/100)

  cols_fill <- c("1" = "#DAA520",
                 "2" = "#b3ff1a",
                 "3" = "#389318",
                 "4" = "#873e23",
                 "5" = "#b3002d",
                 "6" = "#70543e",
                 "7" = 'grey',
                 "8" = "#42c9bc")
plot <- ggplot(NewData)+
  geom_ribbon(aes(x = res, y = fit, ymax = fit+se, ymin = fit-se, 
                  color = true_fct, group = true_fct, fill = true_fct
                  ), alpha = 0.2)+
  geom_line(aes(x = res, y = fit,
                color = true_fct, group = true_fct
                ))+
  # ylim(0,1)+
      scale_fill_manual(name = NULL,
                        values = cols_fill,
                        labels = c("Green algae", "Seagrass", "Brown algae", "Red algae", "Sediment", "Sunglint","Water")) +
    scale_color_manual(name = NULL,
                      values = cols_fill,
                      labels = c("Green algae", "Seagrass", "Brown algae", "Red algae", "Sediment", "Sunglint","Water"))+
  theme_Bede()+
  xlab("Resolution (m)") +
  ylab("Agreement")+
  theme(axis.title = element_text(size = 20),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.key.size = unit(1, "cm"))

ggsave("Output/plot/resolution_plot.png", plot, bg = "white", height = 803*3, width = 1184*3, units = "px")

#### Area

test <-  validation %>% 
  lazy_dt() %>% 
  pivot_longer(c(true, pred), names_to = "source", values_to = "class") %>% 
  group_by(res,class,source) %>% 
  summarise(class_area = n()*(res*res)/(100*100)) %>% 
  as_tibble() %>% 
  pivot_wider(names_from = source, values_from = class_area) %>% 
  mutate(agreement = (1-abs(((pred-true)/(pred+true))))*100,
         # agreement1 = abs(pred-true)/max(pred,true)
         class = as.factor(class))

test %>%
  dplyr::filter(class != 1) %>%
  ggplot(aes(x = res, y =agreement))+
  geom_line(aes(color = class, group = class))

# test %>%
#   pivot_longer(c(pred,true), names_to = "dataset", values_to = "values") %>% 
#   dplyr::filter(class != 1) %>%
#     ggplot(aes(x = res, y = values, color = dataset, group = dataset))+
#   geom_line()+
#   facet_wrap(~ class, scales = "free")

glm_res_area <- glm(agreement ~ res * class , data = test)

summary(glm_res_area)

Utilities.Package::Model_Check(glm_res_area)

NewData<-expand_grid(res=unique(test$res),
                    class=unique(test$class))

pred <- predict(glm_res_area, NewData, se.fit = T)

NewData <- NewData %>% 
  mutate(fit = pred$fit,
         se = pred$se.fit)

NewData %>% 
  dplyr::filter(class != 1) %>% 
ggplot()+
  geom_ribbon(aes(x = res, y = fit, ymax = fit+se, ymin = fit-se, 
                  color = class, group = class, fill = class
                  ), alpha = 0.2)+
  geom_line(aes(x = res, y = fit,
                color = class, group = class
                ))+
  # ylim(0,1)+
      scale_fill_manual(name = NULL,
                        values = cols_fill,
                        labels = c("Green algae", "Seagrass", "Brown algae", "Red algae", "Sediment", "Sunglint","Water")) +
    scale_color_manual(name = NULL,
                      values = cols_fill,
                      labels = c("Green algae", "Seagrass", "Brown algae", "Red algae", "Sediment", "Sunglint","Water"))+
  theme_Bede()+
  xlab("Resolution (m)") +
  ylab("Agreement")+
  theme(axis.title = element_text(size = 20),
        axis.text.x = element_text(size = 15),
        axis.text.y = element_text(size = 15),
        legend.text = element_text(size = 15),
        legend.key.size = unit(1, "cm"))
